Nick now operates a store in Marketview Mall which
      has poor lighting, blasts black metal and sells jeans. Only one pair of
      jeans is available to purchase at a time, though there are more
      stored in the back. If a size is out that you don't want, you must wait
      for someone else to purchase the jeans. Nick's only employee, Hank, sits
      in a chair and stares at people angrily until someone makes a purchase,
      at which point he replaces the jeans with the same model of a random
      size. In order to prevent customers' waiting infinitely for an
      unavailable size, Hank will switch the jeans for a different size pair if
      no one has bought them after a period of three seconds.

\begin{lstlisting}
public class NicksHeavyThreads
{
	// jeans' size [1-5], or 0 when none on display
    private static int awesomeJeans = 0;
    
	// keep this updated as customers arrive and leave
    private static int customers = 0;
    
    private static MeanWorker hank = new MeanWorker();
    
    public static void main( String[] args )
    {
        for( int i = 0; i < 10; ++i )
        {
            ( new LameCustomer() ).start();
        }

		// wait one second before introducing Hank
        try { Thread.sleep( 1000 ); }
        catch( InterruptedException pleaseDont ) {}
        hank.start();
    }
    
    private static class LameCustomer extends Thread
    {
        // ( implementation omitted )
    }
    
    private static class MeanWorker extends Thread
    {
        // ( implementation omitted )
    }
}
\end{lstlisting}
\scriptsize
(Questions may be found on the next page. You may answer them in the space allotted here, or on the following page.)
\normalsize

    \begin{enumerate}

	\pagebreak

    \item Complete the implementation of the \texttt{LameCustomer} class: Each instance
          must choose a jeans size and wait for it to be available, update the
          jeans to indicate that they have been taken, print the message
          ``Customer: I got my size \emph{size} jeans!" and inform all threads
          that the jeans selection has changed. \\
		  \textit{(Hint: Remember to keep an
          accurate count of how many customers are in the shop.)}

\begin{answer}
\begin{lstlisting}
static class LameCustomer extends Thread {
    public void run() {
        synchronized( hank ) {
            ++customers;
			// pick size
            int desiredSize = (int) ( Math.random()*(5) ) + 1;
			// wait for pair
            while( awesomeJeans != desiredSize ) {
                try { hank.wait(); }
                catch( InterruptedException pleaseDont ) {}
            }
            System.out.println( "Customer: I got my size "
                + awesomeJeans + " jeans!" );
            awesomeJeans = 0; // take the jeans
            hank.notifyAll(); // inform everyone they are gone
            --customers;
        }
    }
}
\end{lstlisting}
\end{answer}

    \item Now implement the \texttt{MeanWorker} class, which should choose a size and
          stock a pair of jeans of that size, print the message ``Hank: I
          grumpily restocked with size \emph{size}," and inform all threads
          that the selection has changed. It should then wait until someone has
          taken the jeans or until three seconds have elapsed, whichever comes
          first. These steps should be repeated until all customers have left
          the store.

\begin{answer}
\begin{lstlisting}
static class MeanWorker extends Thread {
    public synchronized void run() {
        do {
			// new size
            awesomeJeans = (int) ( Math.random()*(5) ) + 1;
            System.out.println( "Hank: I grumpily restocked "
                + "with jeans of size " + awesomeJeans );
            notifyAll(); // inform customers of the restocking
            try { wait( 3000 ); } // let people shop
            catch( InterruptedException pleaseDont ) {}
        }
        while( customers > 0 );
    }
}
\end{lstlisting}
\end{answer}
    \end{enumerate}