\begin{enumerate}
\item The following program is intended to read a text file and outputs (as binary data) the number of characters (including new lines) in each line to a file. However, there is a problem with the code the way it is currently written. Find the problem and explain how to fix it. Note: The code is compiled and linked using GCC.

\lstinputlisting{code/getline_binary_io.c}

\begin{answer}
The signature of \texttt{getline} is \texttt{ssize\_t getline(char** lineptr, size\_t* n, FILE* stream)}. If \texttt{*lineptr} is \texttt{NULL} and \texttt{*n} is 0, then \texttt{getline} will dynamically allocate a C string containing the next line of \texttt{stream} (including the new line character), and set \texttt{*lineptr} to this C string, and \texttt{*n} to the size of this new stream. It returns the size of the string read. If the \texttt{*lineptr} given to the function is not \texttt{NULL}, then it expects \texttt{*n} to be the length of \texttt{*lineptr}, and it will attempt to use \texttt{*lineptr} to hold the next line if it can fit, otherwise it will use \texttt{realloc} to allocate enough space for it to fit. This program does not use \texttt{getline} correctly, and causes a type error when compiling. See the answer of the next question for how to fix this problem.
\end{answer}

\item Rewrite the code so that it outputs the data using binary streams instead.
\end{enumerate}

\begin{answer}
\lstinputlisting[basicstyle=\small]{code/getline_binary_io_correct.c}
\end{answer}
